---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Show any code that is needed to  
1.Load the data (i.e. read.csv())
```{r echo=TRUE}
setwd("C:/coursera.datascience/04.repeatableresearch.week02")
sourcefile <- "activity.csv"
sourcedata <- read.csv(sourcefile)
```
2. Process/transform the data (if necessary) into a format suitable for your analysis  
Nothing special required.  

## What is mean total number of steps taken per day?
For this part of the assignment, you can ignore the missing values in the dataset.  
```{r echo=TRUE}
completedata <- na.omit(sourcedata)
```

1. Calculate the total number of steps taken per day  
```{r echo=TRUE eval=TRUE}
workdata <- completedata[1:2]
stepsPerDay <- aggregate(steps ~ date, workdata, sum)
stepsPerDay
```

2. Make a histogram of the total number of steps taken each day  
```{r echo=TRUE eval=TRUE}
hist(stepsPerDay$steps, main="Distribution of steps per day", xlab="Steps per day")
```

3.Calculate and report the mean and median of the total number of steps taken per day  
```{r echo=TRUE eval=TRUE}
mean(stepsPerDay[,2])
median(stepsPerDay[,2])
```
Those values fall in the  highest block in the histogram. This confirms that our calculation makes sense.  

## What is the average daily activity pattern?
1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).  
```{r echo=TRUE eval=TRUE}
fiveMinuteAverage <- aggregate(steps ~ interval, completedata, mean)
plot(fiveMinuteAverage$interval, fiveMinuteAverage$steps, type="l", xlab="5-minute interval", ylab="Average number of steps over all days")
```
2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?  
```{r echo=TRUE eval=TRUE}
fiveMinuteAverage[fiveMinuteAverage$steps == max(fiveMinuteAverage$steps),]
```
This is again confirmed by the graph.  

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
