---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Show any code that is needed to  
1.Load the data (i.e. read.csv())
```{r echo=TRUE}
setwd("C:/coursera.datascience/04.repeatableresearch.week02")
sourcefile <- "activity.csv"
sourcedata <- read.csv(sourcefile)
```
2. Process/transform the data (if necessary) into a format suitable for your analysis  
Nothing special required.  

## What is mean total number of steps taken per day?
For this part of the assignment, you can ignore the missing values in the dataset.  
```{r echo=TRUE}
completedata <- na.omit(sourcedata)
```

1. Calculate the total number of steps taken per day  
```{r echo=TRUE eval=TRUE}
workdata <- completedata[1:2]
stepsPerDay <- aggregate(steps ~ date, workdata, sum)
stepsPerDay
```

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
